---
title: "R Coding Exercise"
---

```{r}
#Load the dslabs package to get the gapminder dataset and other necessary packages
library(dslabs)
library(dplyr)
library(ggplot2)
library(tidyr)
```

```{r}
#Exploratory analysis of the gapminder dataset

#Use the help method to get a data description
help("gapminder")

#Get an overview of data structure
str(gapminder)

#Get a data summary
summary(gapminder)

#Determine the type of object that gapminder is
class(gapminder)
```

```{r}
#Data processing

#Create a subset of the data just containing countries from Africa
africadata <- gapminder %>%
  filter(continent == "Africa")

#Examine the structure of the Africa-specific data
str(africadata)
summary(africadata)

#Create a new subset of the africa data only containing data on infant mortality and life expectancy
africa1 <- africadata %>%
  select(c('infant_mortality', 'life_expectancy'))

#Examine the structure of the first subset
str(africa1)
summary(africa1)

#Create a new subset of the africa data only containing data on population and life expectancy
africa2 <- africadata %>%
  select(c('population', 'life_expectancy'))

#Examine the structure of the second subset
str(africa2)
summary(africa2)
```

```{r}
#Data plotting

#Plot life expectancy as a function of infant mortality
ggplot(africa1, aes(x = infant_mortality, y = life_expectancy)) + 
  geom_point() + #Make it a scatter plot
  labs(
    title = "Life expectancy vs. infant mortality in Africa",
    x = "Infant Mortality (deaths per 1,000)",
    y = "Life Expectancy (years)"
  ) + #Add plot title and axis labels
  theme_minimal() #Clean theme
  
#Plot life expectancy as a function of population
ggplot(africa2, aes(x = log(population), y = life_expectancy)) +
  geom_point() + #Make it a scatter plot
  labs(
    title = "Life expectancy vs. log population in Africa",
    x = "Log Population",
    y = "Life Expectancy (years)"
  ) + #Add plot title and axis labels
  theme_minimal() #Clean theme
```

The plot of life expectancy vs. infant mortality shows a slight negative association. The plot of life expectancy vs. log population shows a strong positive association between population size and life expectancy. The data displayed in both graphs, especially in the second graph shows individualized trends in life expectancy. This is likely because data for every year included in the dataset is included in the plot, showing individual trends within countries as well as overall trends in the continent.

```{r}
#Further data processing

#Find the year with the most data (non-NA values) for infant mortality
most_data_year <- africadata %>% #Make a copy of the africa data
  group_by(year) %>% #Group by year
  summarize(non_na_count = sum(!is.na(infant_mortality))) %>% #Count non-NA values, summarize in new column
  arrange(desc(non_na_count)) #Sort non_na_count in descending order

#Display year with most data
print(most_data_year)
```

By inspecting the number of available infant mortality data points by year, complete data is available from 1982-2015. Years before and after that period have at least one missing data point. For further analysis, I will focus on the year 2000.

```{r}
#Filter africa data to only include the year 2000
africadata_2000 <- africadata %>%
  filter(year == 2000)

#Plot life expectancy as a function of infant mortality in the year 2000
ggplot(africadata_2000, aes(x = infant_mortality, y = life_expectancy)) + 
  geom_point() + #Make it a scatter plot
  labs(
    title = "Life expectancy vs. infant mortality in Africa in 2000",
    x = "Infant Mortality (deaths per 1,000)",
    y = "Life Expectancy (years)"
  ) + #Add plot title and axis labels
  theme_minimal() #Clean theme
  
#Plot life expectancy as a function of population
ggplot(africadata_2000, aes(x = log(population), y = life_expectancy)) +
  geom_point() + #Make it a scatter plot
  labs(
    title = "Life expectancy vs. log population in Africa in 2000",
    x = "Log Population",
    y = "Life Expectancy (years)"
  ) + #Add plot title and axis labels
  theme_minimal() #Clean theme
```

After plotting only the data from 2000, a negative association was still seen between life expectancy and infant mortality. However, no clear association can be seen between population size and infant mortality using only data from 2000.

```{r}
#Regression modeling

#Fit a linear model predicting life expectancy from infant mortality with data from 2000
lm1 <- lm(life_expectancy ~ infant_mortality, africadata_2000)
summary(lm1) #Present summary statistics of the linear model

#Fit a linear model predicting life expectancy from log population with data from 2000
lm2 <- lm(life_expectancy ~ log(population), africadata_2000)
summary(lm2) #Present summary statistics of the linear model
```

From the linear model fitting life expectancy from infant mortality, I can conclude that infant mortality is a significant predictor of life expectancy (t = -6.594, p \< 0.001). From the linear model fitting life expectancy from log population size, I *cannot* conclude that population size is a significant predictor of life expectancy (t = -0.719, p = 0.476).
