---
title: "tidytuesday-exercise"
author: "Connor Norris"
editor: visual
---

# Tidy Tuesday: Timely and Effective Care by US State

For this exercise, I will analyze the tidytuesday dataset for the week of April 8, 2025.

```{r, message = FALSE}
#Load required packages
library(tidytuesdayR)
library(tidyverse)
library(tidymodels)
library(ggplot2)
library(here)
library(car)
library(ranger)
library(xgboost)
library(broom)
library(vip)
library(corrplot)
```

```{r, message = FALSE}
#Load data
tuesdata <- tidytuesdayR::tt_load('2025-04-08')
care_state <- tuesdata$care_state

#Save data in repo
write_csv(x = care_state, "care_state.csv")
```

## Data Wrangling

First, I'll get an overhead view of the data

```{r}
df <- care_state
#Data structure
str(df)

#Summary stats of each variable
summary(df)

#Get unique measure names
unique(df$measure_name)

#Skim
skimr::skim(df)

summary(filter(df, measure_id == "SAFE_USE_OF_OPIOIDS")$score)
```

The data appears to have a number of different measures related to healthcare, and some of the measures are represented multiple times in individual strata (e.g. the average time patients spend in the ER and the amount of time patients spend in the ER (high)). For this analysis, I'd like to do predictive modeling to see if other outcome scores can predict the percentage of healthcare personnel who are up to date with COVID-19 vaccinations.

```{r}
#Keep only relevant measures
relevant_measures <- c("HCP_COVID_19", "IMM_3", "OP_18b", "OP_18c", "OP_22", "OP_23", "OP_29", "OP_31", "SEP_1", "SAFE_USE_OF_OPIOIDS")

#Filter the data to only include measures of interest
df <- df %>%
  filter(measure_id %in% relevant_measures)

#Transform to wide format
df_wide <- df %>%
  select(state, measure_id, score) %>% #Keep only relevant columns
  pivot_wider(names_from = measure_id, values_from = score)

#Inspect the new data
skimr::skim(df_wide)
head(df_wide)
```

Now my data is in a much more suitable format for modelling, with each state getting one row, and the scores for each measure of interest being in their own columns. However, the measure OP-31 has much more missing data than the rest, so I will remove this measure.

```{r}
#Remove OP-31, then remove rows with NA values
df_wide <- df_wide %>%
  select(-OP_31) %>%
  drop_na()

#Save clean data
write_csv(df_wide, "care_state_clean.csv")
```

In handling missing values, 4 states/territories were dropped from the data, leaving 52 states/territories.

## Exploratory Data Analysis

```{r}
#Convert back to long format for facet wrapping
df_long <- pivot_longer(df_wide, cols = -state, names_to = "Measure", values_to = "Value")

#Create faceted histogram plot
ggplot(df_long, aes(x = Value)) +
  geom_histogram(binwidth = 2.5, fill = "royalblue", color = "black") +
  facet_wrap(~Measure, scales = "free") + #Free scales for each facet
  labs(
    title = "Histograms of Measures",
    x = "Value",
    y = "Frequency"
  ) + 
  theme_minimal()

#Create boxplots for each measure
ggplot(df_long, aes(x = Measure, y = Value)) +
  geom_boxplot(fill = "royalblue", color = "black") +
  facet_wrap(~Measure, scales = "free") +
  labs(
    title = "Boxplots of Measures",
    x = "",
    y = "Value"
  ) +
  theme_minimal()

#Correlation analysis
cor_matrix <- cor(df_wide[, -1], use = "complete.obs")
cor_matrix
corrplot(cor_matrix)

#Scatterplot matrix
scatterplotMatrix(df_wide[, -1])
```

Based on the histograms and boxplots, it seems that all of the scores for each measure have unique ranges. It also appears that OP_18b and OP_18c are strongly associated, which will need to be handled in the modelling.

## Modeling

```{r}
#Set random seed
rngseed = 1234
set.seed(1234)

#Data splitting (80% training, 20% testing)
data_split <- initial_split(df_wide, prop = 0.8)
train <- training(data_split)
test <- testing(data_split)

#Cross-validation: 5-fold CV
set.seed(rngseed)
cv_folds <- vfold_cv(train, v = 5)

#Define models
#Linear regression
lm_model <- linear_reg() %>%
  set_engine("lm")

#Random forest
rf_model <- rand_forest(mtry = 3, trees = 500, min_n = 5) %>%
  set_engine("ranger", importance = "permutation") %>%
  set_mode("regression")

#XGBoost
xgb_model <- boost_tree(trees = 1000, tree_depth = 6, learn_rate = 0.1) %>%
  set_engine("xgboost", importance = "permutation") %>%
  set_mode("regression")

#Define a recipe for modeling
model_recipe <- recipe(HCP_COVID_19 ~ ., data = train[, -1]) %>%
  step_normalize(all_numeric_predictors()) %>% #Standardize numeric predictors 
  step_interact(terms = ~ OP_18b:OP_18c) #Add interaction term

#Create workflows for each model
lm_workflow <- workflow() %>%
  add_model(lm_model) %>%
  add_recipe(model_recipe)

rf_workflow <- workflow() %>%
  add_model(rf_model) %>%
  add_recipe(model_recipe)

xgb_workflow <- workflow() %>%
  add_model(xgb_model) %>%
  add_recipe(model_recipe)

#Fit models using cross-validation
set.seed(rngseed)
lm_results <- fit_resamples(lm_workflow, resamples = cv_folds, metrics = metric_set(rmse, rsq))
rf_results <- fit_resamples(rf_workflow, resamples = cv_folds, metrics = metric_set(rmse, rsq))
xgb_results <- fit_resamples(xgb_workflow, resamples = cv_folds, metrics = metric_set(rmse, rsq))

#Collect metrics
lm_metrics <- collect_metrics(lm_results)
rf_metrics <- collect_metrics(rf_results)
xgb_metrics <- collect_metrics(xgb_results)

#Compare metrics
lm_metrics
rf_metrics
xgb_metrics
```

## Variable Selection

```{r}
set.seed(rngseed)
#Use a lasso for the linear model
#Define a lasso model
lasso_model <- linear_reg(penalty = tune(), mixture = 1) %>%
  set_engine("glmnet")

#Create a workflow for the lasso
lasso_workflow <- workflow() %>%
  add_recipe(model_recipe) %>%
  add_model(lasso_model)

#Get results
lasso_results <- tune_grid(
  lasso_workflow,
  resamples = cv_folds,
  grid = 10,
  metrics = metric_set(rmse, rsq)
)

#Select the best penalty
best_lasso <- select_best(lasso_results, metric = "rmse")

#Finalize the workflow with the best penalty
final_lasso <- finalize_workflow(lasso_workflow, best_lasso)

#Fit the finalized model
final_lm_fit <- fit(final_lasso, data = train)

#Extract coefficients
tidy(final_lm_fit$fit$fit) %>% filter(estimate != 0)
```

```{r}
#Fit random forest
rf_fit <- fit(rf_workflow, data = train)

#Extract variable importance
vip(rf_fit$fit$fit)

#Fit xgb
xgb_fit <- fit(xgb_workflow, data = train)

#Extract variable importance
vip(xgb_fit$fit$fit)
```

Each model had different significant predictors, and I will re-fit each model with those predictors.

## Re-evaluating models using the selected variables

```{r}
#Define new recipe for the linear regression
lm_recipe <- recipe(HCP_COVID_19 ~ IMM_3 + OP_18c + OP_22 + OP_23 + OP_29 + SAFE_USE_OF_OPIOIDS+ SEP_1, data = train[, -1]) %>%
  step_normalize(all_numeric_predictors()) #Standardize numeric predictors 

#Define new recipe for the random forest
rf_recipe <- recipe(HCP_COVID_19 ~ OP_18b + OP_22 + OP_18c + SEP_1 + OP_29 + OP_23, data = train[, -1]) %>%
  step_normalize(all_numeric_predictors()) #Standardize numeric predictors

#Define new recipe for the XGBoost
xgb_recipe <- recipe(HCP_COVID_19 ~ OP_18b + IMM_3 + OP_18c + OP_29 + OP_22 + SAFE_USE_OF_OPIOIDS + OP_23, data = train[,-1]) %>%
  step_normalize(all_numeric_predictors())

#Reset workflows
lm_workflow <- workflow() %>%
  add_model(lm_model) %>%
  add_recipe(lm_recipe)

rf_workflow <- workflow() %>%
  add_model(rf_model) %>%
  add_recipe(rf_recipe)

xgb_workflow <- workflow() %>%
  add_model(xgb_model) %>%
  add_recipe(xgb_recipe)

#Fit models using cross-validation
set.seed(rngseed)
lm_results <- fit_resamples(lm_workflow, resamples = cv_folds, metrics = metric_set(rmse, rsq))
rf_results <- fit_resamples(rf_workflow, resamples = cv_folds, metrics = metric_set(rmse, rsq))
xgb_results <- fit_resamples(xgb_workflow, resamples = cv_folds, metrics = metric_set(rmse, rsq))

#Collect metrics
lm_metrics <- collect_metrics(lm_results)
rf_metrics <- collect_metrics(rf_results)
xgb_metrics <- collect_metrics(xgb_results)

#Compare metrics
lm_metrics
rf_metrics
xgb_metrics
```

```{r}
#Compare to a null model
#Define null model
null_mod <- null_model() %>%
  set_engine("parsnip") %>%
  set_mode("regression")

#Make a recipe for the null model
null_recipe <- recipe(HCP_COVID_19 ~ 1, data = train[, -1])

#Create a workflow for the null model
null_workflow <- workflow() %>%
  add_recipe(null_recipe) %>%
  add_model(null_mod)

#Fit the null model with cross-validation
set.seed(rngseed)
null_results <- fit_resamples(
  null_workflow,
  resamples = cv_folds,
  metrics = metric_set(rmse, rsq)
)

#Show metrics
collect_metrics(null_results)
```

The RMSE of the null model was lower than all of the other models used here, so none of the models I've fit are very informative. However, of the three, the random forest performed the best, as it had the lowest RMSE and the highest R-squared. I will try to fit the testing data on the random forest.

```{r}

```

```{r}
#Predict testing data using the random forest model
test_preds <- predict(rf_fit, new_data = test) %>%
  bind_cols(test)

#Evaluate performance on testing data
test_metrics <- test_preds %>%
  metrics(truth = HCP_COVID_19, estimate = .pred)

test_metrics
```

The RMSE was lower using the testing data, but the R-squared was almost zero. Based on this analysis, none of the predictors included in this data set were significant predictors of the percentage of healthcare personnel who are up to date with COVID-19 vaccinations.
